A Study of Common Bug Fix Patterns in Rust

### 2022-11-23 discussion.tex
applied txt fixes for results and consecutive sections
added changes

In our work there are three main threats to internal validity: (1) We use a weighting scheme which is based on a heuristic to give more values to tree elements closer to the leaves. In addition to that, we also manually adjusted the weights of borrow checker related elements to respect ownership framework's role in Rust, though we recorded their results separate from general results. As this approach is based on an intuition, it does not guarantee achieving optimal clusters. (2) The collected commits are the ones that fix a bug in a previously pushed commit. We do not consider the bugs that occur during a commit and also fixed during the same commit. (3) Our data modelling approach is based on the frequency of observed program elements in ASTs, and we use DBSCAN as our clustering algorithm. Other program data modelling methods and clustering algorithms (e.g.SLINK) might output new clusters that our pipeline is unable to find.

>>>>

In our work there are three main threats to internal validity: (1) We use a weighting scheme which is based on a heuristic to give more values to tree elements closer to the leaves. In addition to that, we also manually adjusted the weights of borrow checker related elements to respect ownership framework's role in Rust, though we recorded their results separate from general results. As this approach is based on an intuition, it does not guarantee achieving optimal clusters. (2) The collected commits are the ones that fix a bug in a previously pushed commit. We do not consider the bugs that occur during a commit and also fixed during the same commit. (3) Our code embedding approach is based on the frequency of observed program elements in ASTs, and we use DBSCAN as our clustering algorithm. Other program code embedding methods and clustering algorithms (e.g. SLINK) might output new clusters that our pipeline is unable to find.

### 2022-11-23 introduction.tex
added changes

Bugs have long been an unavoidable part of computer programming, predating Grace Hopper's literal debugging session in 1947; even today, bug-free software remains an aspirational aim for software engineering research. One step towards that aim has been studying common bug patterns (and their corresponding fixes) across multiple software systems. Research efforts categorizing common bug patterns date back to at least 1975, where Endres tried to categorize bug patterns in operating system implementations~\citep{endres1975analysis}.

In this work, with the help of our tool, Ruxanne, we empirically analyze the most popular Rust projects in GitHub. Our goal is to discover common bug patterns that afflict these projects. Although we are motivated by a desire to develop automated program repair tools, we believe these insights can also be of broader service. Automated program repair~\citep{goues2019automated,liu2018survey} tools try to find bug locations in the source code (using a fault localization module) and then generate patches to fix them, possibly without human intervention, so that the modified program meets desired specifications (implicit or explicit). A powerful automated program repair system has many useful applications and could significantly simplify debugging, thus reducing software development cost~\citep{le2012systematic}. A set of common bug fix patterns can be a useful input to an APR system: it tells how to fix recurring buggy structures, thus reducing the search space for patch generation~\citep{jeffrey2009bugfix}. 
Yet, to the best of our knowledge, no research has sought to identify the most pervasive bug fix patterns within Rust programs. In this project, we introduce Ruxanne, a tool for analyzing and extracting fix patterns in Rust. Ruxanne uses a novel embedding of Rust code into fixed-sized vectors. Through Ruxanne, we mined the top 18 most-starred Rust projects in Github to discover the most common bug fix patterns committed to their repositories. We analyzed 87,726 code changes drawn from 57,214 commits across these 18 projects. After clustering the code changes, and conducting a manual analysis, we discovered 20 groups of cross-project bug fix patterns, which we categorize as (1) general patterns and (2) borrow-checker-related patterns. We describe each of these patterns and their implications to automated program repair.

@@

Many studies have investigated detecting and categorizing bug fix patterns in general~\citep{islam2021changes,madeiral2018towards,pan2009toward}, or for a specific programming language~\citep{yang2022mining,hanam2016discovering,campos2019discovering}. In this work, we study common bug patterns that appear in Rust projects. Rust is a relatively new programming language~\citep{klabnik2019rust} which allows programmers to write programs that have low-level control over resources, while still ensuring high-level safety guarantees (for programs written in safe Rust). What distinguishes Rust from the other systems programming languages, e.g. C/C++, is its ownership framework, which is an integral part of Rust. The ownership framework keeps track of references to different memory locations, and will free them once they go out of scope. If a program passes the compiler’s checks, it meets several safety properties by construction; e.g. it has no dangling references or double-frees. Such properties had to be manually ensured (or not!) in languages like C/C++. Moreover, not only does ownership remove the need for a garbage collector in the runtime environment, it frees the programmer from the need to explicitly do certain safety-related tasks~\citep{qin2020understanding}. Due to the memory and thread safety guarantees that Rust provides, many software companies have adopted Rust in production. For example, Dropbox decided to migrate to Rust for developing their storage system, since they were not completely satisfied with the features provided by AWS’s S3 or Golang~\citep{dropbox}.

@@

\begin{itemize}
    \item \textbf{RQ1. Does our data modelling approach capture the most important aspects of bug-fixing program changes?} 
    \item \textbf{RQ2. What are the general fix patterns in Rust, and how often do they apply?} 
    \item \textbf{RQ3. What are borrow-checker related fix patterns in Rust, and how often do they apply?}
\end{itemize}

@@

\begin{itemize}
    \item To the best of our knowledge, this work is the first to exploits a code analysis pipeline to automatically mine Rust open source projects and extract common cross-project bug fix patterns.
    \item We propose a novel method to encode key AST information in fixed size datapoints using a semi-automatically derived weighting scheme.
    \item To respect the key role of the borrow checker in Rust, we report and discuss bug fix patterns that are related to the ownership framework in Rust separately from general patterns.
\end{itemize}

>>>>

Bugs have long been an unavoidable part of computer programming, predating Grace Hopper's literal debugging session in 1947; even today, bug-free software remains an aspirational aim for software engineering research. One step towards that aim has been studying common bug patterns (and their corresponding fixes) across multiple software systems. Research efforts categorizing common bug patterns date back to at least 1975, where \cite{endres1975analysis} tried to categorize bug patterns in operating system implementations.

In this work, with the help of our tool, Ruxanne, we empirically analyze the most popular Rust projects in GitHub. Our goal is to discover common bug patterns that afflict these projects. Although we are motivated by a desire to develop automated program repair tools, we believe these insights can also be of broader service. Automated program repair~\citep{goues2019automated,liu2018survey} tools try to find bug locations in the source code (using a fault localization module) and then generate patches to fix them, possibly without human intervention, so that the modified program meets desired specifications (implicit or explicit). A powerful automated program repair system has many useful applications and could significantly simplify debugging, thus reducing software development cost~\citep{le2012systematic}. A set of common bug fix patterns can be a useful input to an APR system: it tells how to fix recurring buggy structures, thus reducing the search space for patch generation~\citep{jeffrey2009bugfix}. 
Yet, to the best of our knowledge, no research has sought to identify the most pervasive bug fix patterns in Rust programs. In this project, we introduce Ruxanne, a tool for analyzing and extracting fix patterns in Rust. Ruxanne uses a novel embedding of Rust code into fixed-sized vectors. Through Ruxanne, we mined the top 18 most-starred Rust projects in Github to discover the most common bug fix patterns committed to their repositories. We analyzed 87,726 code changes drawn from 57,214 commits across these 18 projects. After clustering the code changes, and conducting a manual analysis, we discovered 20 groups of cross-project bug fix patterns, which we categorize as (1) general patterns and (2) borrow-checker-related patterns. We describe each of these patterns and their implications to automated program repair.

@@

Many studies have investigated detecting and categorizing bug fix patterns in general~\citep{islam2021changes,madeiral2018towards,pan2009toward}, or for a specific programming language~\citep{yang2022mining,hanam2016discovering,campos2019discovering}. In this work, we study common bug patterns that appear in Rust projects. Rust is a relatively new programming language~\citep{klabnik2019rust} which allows programmers to write programs that have low-level control over resources, while still ensuring high-level safety guarantees (for programs written in safe Rust). What distinguishes Rust from the other systems programming languages, e.g. C/C++, is its ownership framework, which is an integral part of Rust. The ownership framework keeps track of references to different memory locations, and will free them once they go out of scope. If a program passes the compiler’s checks, it meets several safety properties by construction; e.g. it has no dangling references or double-frees. Such properties had to be manually ensured (or not!) in languages like C/C++. Moreover, not only does ownership remove the need for a garbage collector in the runtime environment, it frees the programmer of the responsibility to ensure certain safety-related properties~\citep{qin2020understanding}. Due to the memory and thread safety guarantees that Rust provides, many software companies have adopted Rust in production. For example, Dropbox decided to migrate to Rust for developing their storage system, since they were not completely satisfied with the features provided by AWS’s S3 or Golang~\citep{dropbox}.

@@

\begin{itemize}
    \item \textbf{RQ1. Does our code embedding approach capture the most important aspects of bug-fixing program changes?} 
    \item \textbf{RQ2. What are the general fix patterns in Rust, and how often do they apply?} 
    \item \textbf{RQ3. What are borrow-checker related fix patterns in Rust, and how often do they apply?}
\end{itemize}

@@

\begin{itemize}
    \item To the best of our knowledge, this work is the first to exploit a code analysis pipeline to automatically mine Rust open source projects and extract common cross-project bug fix patterns.
    \item This work proposes a novel method to encode key AST information in fixed size datapoints using a semi-automatically derived weighting scheme.
    \item Keeping in mind the key role of the borrow checker in Rust, this work specifically identifies bug fix patterns that are related to Rust's ownership framework in addition to general bug fix patterns for Rust.
\end{itemize}

### 2022-11-23 main.tex
added changes

\title{A Study on Common Bug Fix Patterns in Rust%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}
\subtitle{}

@@

\institute{F. Author \at
              first address \\
              Tel.: +123-45-678910\\
              Fax: +123-45-678910\\
              \email{fauthor@example.com}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           S. Author \at
              second address
}

@@

\begin{abstract}
    Rust is a relatively new programming language which allows programmers to write programs that have low-level control over resources, while still ensuring high-level safety guarantees (for programs written in safe Rust). Rust's ownership framework enables programs to meet these two seemingly-contradictory goals. The Rust compiler's Borrow-Checker component enforces the ownership framework requirements which ensure Rust's safety guarantees.

@@

Yet, to the best of our knowledge, no research has sought to identify the most pervasive bug fix patterns within Rust programs. In this project, we introduce Ruxanne, a tool for analyzing and extracting fix patterns in Rust. Ruxanne uses a novel embedding of Rust code into fixed-sized vectors. Through Ruxanne, we mined the top 18 most-starred Rust projects in Github to discover the most common bug fix patterns committed to their repositories. We analyzed 87,726 code changes drawn from 57,214 commits across these 18 projects. After clustering the code changes, and conducting a manual analysis, we discovered 20 groups of cross-project bug fix patterns, which we categorize as (1) general patterns and (2) borrow-checker-related patterns. We describe each of these patterns and their implications to automated program repair.
\todo[inline]{Do we need this last sentence?}
\keywords{Bug patterns \and Pattern mining \and Bug fix changes \and Rust}
% \PACS{PACS code1 \and PACS code2 \and more}
\subclass{68N01 \and 68N15} %% Computer Science : General and Programming Languages
\end{abstract}

>>>>

\title{A Study of Common Bug Fix Patterns in Rust%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}
\subtitle{}

@@

\institute{Mohammad Robati Shirzad \at
              200 University Ave W, Waterloo, ON N2L 3G1 \\
              \email{mrobatis@uwaterloo.ca}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           Patrick Lam \at
                200 University Ave W, Waterloo, ON N2L 3G1 \\
                \email{patrick.lam@uwaterloo.ca}
}

@@

\begin{abstract}
    Rust is a relatively new programming language which allows programmers to write programs that have low-level control over resources while still ensuring high-level safety guarantees (for programs written in safe Rust). Rust's ownership framework enables programs to meet these two seemingly-contradictory goals. The Rust compiler's Borrow-Checker component enforces the ownership framework requirements that ensure Rust's safety guarantees.

@@

Yet, to the best of our knowledge, no research has sought to identify the most pervasive bug fix patterns within Rust programs. In this project, we introduce Ruxanne, a tool for analyzing and extracting fix patterns in Rust. Ruxanne implements a novel embedding of Rust code into fixed-sized vectors. Using Ruxanne, we mined the top 18 most-starred Rust projects in Github to discover the most common bug fix patterns committed to their repositories. We analyzed 87,726 code changes drawn from 57,214 commits across these 18 projects. After clustering the code changes, and conducting a manual analysis, we discovered 20 groups of cross-project bug fix patterns, which we categorize as (1) general patterns and (2) borrow-checker-related patterns. We describe each of these patterns and their implications to automated program repair.
\keywords{Bug patterns \and Pattern mining \and Bug fix changes \and Rust}
% \PACS{PACS code1 \and PACS code2 \and more}
\subclass{68N01 \and 68N15} %% Computer Science : General and Programming Languages
\end{abstract}
